The SMT Competition 2015--2018

### 2018-11-24 2018/report/SMT-COMP.tex
Affiliations
Weights, disagreements, minor

\author{%
  \name{Tjark Weber (competition chair, 2015--2018)}
  \email{tjark.weber@it.uu.se} \\
  \addr Uppsala University \\
  Uppsala \\
  Sweden
  \AND
  \name{Sylvain Conchon (co-organizer, 2015--2016)}
  \email{Sylvain.Conchon@lri.fr} \\
  \addr Affiliation \\
  City \\
  Country
  \AND
  \name{David D\'{e}harbe (co-organizer, 2015--2016)}
  \email{david.deharbe@clearsy.com} \\
  \addr ClearSy Systems Engineering \\
  Aix-en-Provence \\
  France
  \AND
  \name{Matthias Heizmann (co-organizer, 2016--2018)}
  \email{heizmann@informatik.uni-freiburg.de} \\
  \addr Affiliation \\
  City \\
  Country
  \AND
  \name{Aina Niemetz (co-organizer, 2018)}
  \email{niemetz@cs.stanford.edu} \\
  \addr Stanford University\\
  Stanford \\
  USA
  \AND
  \name{Giles Reger (co-organizer, 2017--2018)}
  \email{giles.reger@manchester.ac.uk} \\
  \addr Affiliation \\
  City \\
  Country}

@@

For space reasons, the discussion in this paper is necessarily
limited.  Complete data about past competitions since 2015 (and partly
for earlier competitions) is available from the competition website,
\url{http://www.smtcomp.org}.  This includes detailed results for each
solver/benchmark combination, totaling nearly 6 million data records
(about 1.1~GB in CSV format).

The rest of this paper is structured as follows.  In
Section~\ref{sec:goals}, we present the goals and organization of the
competition.  Section~\ref{sec:smtlib} introduces the SMT-LIB language
and library, and Section~\ref{sec:benchmarks} discusses competition
divisions and benchmarks.  The competition is being run on StarExec,
which is described in Section~\ref{sec:starexec}.
Section~\ref{sec:participants} gives an overview of participating
solvers.  The details of the competition procedure are described in
Section~\ref{sec:procedure}, and selected results are presented and
analyzed in Section~\ref{sec:results}.  We conclude with suggestions
for future SMT Competitions in Section~\ref{sec:conclusions}.

@@

SMT Competitions until~2014 employed a pseudo-random selection process
to further cull the number of benchmarks used in the competition.
However, a 2013 evaluation of SMT-COMP and SMT-LIB found that this
``significantly lessens the ability of a competition to determine
`best' solvers.''~\cite{CSW15} Together with the increased
computational resources available to the competition since its move to
StarExec, this prompted a paradigm shift.  Since 2015, the competition
is evaluating all solvers on all eligible benchmarks. \TODO: benchmark
weights

Another paradigm shift concerns the treatment of benchmarks with
unknown status.  Because of the difficulty to judge solver responses
for such benchmarks, SMT Competitions until 2015 only used benchmarks
whose status was known (i.e., that had been solved previously by
existing SMT solvers).  In 2015, 30,717 non-incremental benchmarks
from the SMT Library were not used for the competition because their
status was unknown.  However, this approach was criticized for
favoring imitation of existing solvers over true innovation.  In~2016,
the SMT Library contained 29,724 non-incremental benchmarks with
unknown status; solver performance on these was evaluated but reported
separately.  Since 2017, all non-incremental benchmarks are eligible
for the competition's main track regardless of status, and they
directly affect the competition results.  Solver responses for such
benchmarks are assumed to be correct, unless there is a disagreement
between two or more (otherwise sound) solvers.  Benchmarks with
disagreements are removed from the competition results (see
Section~\ref{sec:results-disagreements} for further discussion).

@@

Also shown in Figure~\ref{fig:job-pairs} is the total wall-clock time
(summed over all job pairs) for the competition.  This has more than
doubled, from 3.5~years in~2015 to 7.6~years in~2018.  (These are
solver run-times only.  Time that the cluster spent on pre- and
post-processing of benchmarks and solver output or on other auxiliary
tasks is not included.)  The increase in~2016 is largely due to the
inclusion of (non-incremental) benchmarks with unknown status; these
were evaluated with a wall-clock time limit of 10~minutes.  In~2017,
when these benchmarks had become eligible for the main track,
main-track job pairs were first run with a time limit of 10~minutes.
When it became apparent to the organizers that there were sufficient
resources to increase this limit, job pairs that had timed out were
re-run with a time limit of 20~minutes.  (In this way, about 2.9 years
of wall-clock time in 2017 were spent on job pairs that were then
re-run.  This time is not included in Figure~\ref{fig:job-pairs}.)
Table~\ref{table:time-limits} shows the wall-clock time limits that
were imposed on job pairs for each track and year.

@@

\begin{table}
  \caption{Participants (2015--2018)}
  \label{table:participants}
  \centering
  \begin{tabular}{|l|p{11.5cm}|} %TODO: fix width
    \hline
    Solver & Submitters (Affiliations)\\
    \hline
    ABC~\cite{DBLP:conf/cav/BraytonM10} & Valeriy Balabanov
                        (Mentor Graphics)\\
    Alt-Ergo          & Albin Coquereau
                        (CNRS and University of Paris-Sud)\\
    AProVE~\cite{DBLP:conf/cade/GieslBEFFOPSSST14} & Carsten Fuhs
                        (University College London; Birkbeck, University of London)\\
    Boolector~\cite{DBLP:conf/cav/NiemetzPWB18} & Mathias Preiner
                        (Johannes Kepler University)\\
    COLIBRI~\cite{colibri} & Fran\c{c}ois Bobot
                        (French Alternative Energies and Atomic Energy Commission)\\
    Ctrl-Ergo         & Mohamed Iguernlala
                        (CNRS and University of Paris-Sud)\\
    CVC3~\cite{DBLP:conf/cav/BarrettT07} & Kshitij Bansal
                        (New York University)\\
    CVC4~\cite{DBLP:conf/cav/BarrettCDHJKRT11}  & Kshitij Bansal, Clark Barrett, Andres N\"otzli, Duligur Ibeling
                        (New York University; Stanford University)\\
    MapleSTP          & Jimmy Liang
                        (University of Waterloo)\\
    MinkeyRink        & Trevor Hansen
                        (University of Melbourne)\\
    OpenSMT2~\cite{DBLP:conf/sat/HyvarinenMAS16} & Matteo Marescotti, Antti Hyv\"arinen
                        (University of Lugano)\\
    ProB~\cite{doi:10.1002/9781119002727.ch14} & Sebastian Krings
                        (University of D\"usseldorf)\\
    Q3B               & Martin Jon\'a\v{s}
                        (Masaryk University)\\
    raSAT~\cite{rasat} & Tung Vu Xuan
                        (Japan Advanced Institute of Science and Technology)\\
    Redlog~\cite{DBLP:journals/cca/Dolzmann097} & Haniel Barbosa
                        (Inria Nancy)\\
    SMT-RAT~\cite{DBLP:conf/sat/CorziliusKJSA15} & Florian Corzilius, Gereon Kremer
                        (RWTH Aachen University)\\
    SMTInterpol~\cite{DBLP:conf/spin/ChristHN12} & J\"urgen Christ
                        (University of Freiburg)\\
    SPASS-SATT         & Martin Bromberger
                        (Max Planck Institute for Informatics)\\
    STP~\cite{ganesh07} & Mate Soos, Trevor Hansen, Norbert Manthey
                        (National University of Singapore; University of Melbourne; Amazon Web Services)\\
    toysmt             & Masahiro Sakai
                        (Toshiba Corporation)\\
    Vampire~\cite{DBLP:conf/cav/KovacsV13} & Giles Reger
                        (University of Manchester)\\
    veriT~\cite{DBLP:conf/cade/BoutonODF09} & David D\'eharbe, Haniel Barbosa, Hans-J\"org Schurr
                        (Federal University of Rio Grande do Norte; Inria Nancy)\\
    XSat~\cite{DBLP:conf/cav/FuS16} & Martin Velez
                        (University of California, Davis)\\
    Yices2~\cite{Dutertre:cav2014} & Bruno Dutertre, Dejan Jovanovi\'c
                        (SRI International)\\
    Z3~\cite{DBLP:conf/tacas/MouraB08} & Christoph Wintersteiger, Aleksandar Zelji\'c
                        (Microsoft Research; Uppsala University)\\
    \hline
  \end{tabular}
\end{table}

@@

\section{Competition Procedure}
\label{sec:procedure}

\TODO: explain those aspects of the procedure that are required to
understand the results (e.g., scoring)

@@

@@

>>>>

\author{%
  \name{Tjark Weber (competition chair, 2015--2018)}
  \email{tjark.weber@it.uu.se} \\
  \addr Uppsala University \\
  %Uppsala \\
  Sweden
  \AND
  \name{Sylvain Conchon (co-organizer, 2015--2016)}
  \email{Sylvain.Conchon@lri.fr} \\
  \addr Paris-Sud University \\
  France
  \AND
  \name{David D\'{e}harbe (co-organizer, 2015--2016)}
  \email{david.deharbe@clearsy.com} \\
  \addr ClearSy Systems Engineering \\
  %Aix-en-Provence \\
  France
  \AND
  \name{Matthias Heizmann (co-organizer, 2016--2018)}
  \email{heizmann@informatik.uni-freiburg.de} \\
  \addr University of Freiburg \\
  %City \\
  Germany
  \AND
  \name{Aina Niemetz (co-organizer, 2018)}
  \email{niemetz@cs.stanford.edu} \\
  \addr Stanford University\\
  %Stanford \\
  USA
  \AND
  \name{Giles Reger (co-organizer, 2017--2018)}
  \email{giles.reger@manchester.ac.uk} \\
  \addr University of Manchester \\
  UK}

@@

For space reasons, the discussion in this paper is necessarily
limited.  Complete data about past competitions since 2015 (and partly
for earlier competitions) is available from the competition website,
\url{http://www.smtcomp.org}.  This includes detailed results for each
solver/benchmark combination, totaling nearly 6 million data records
(about 1.1~GB in CSV format).  Competition solvers and benchmarks may
be downloaded from StarExec, \url{https://www.starexec.org}.

The rest of this paper is structured as follows.  In
Section~\ref{sec:goals}, we present the goals and organization of the
competition.  Section~\ref{sec:smtlib} introduces the SMT-LIB language
and library, and Section~\ref{sec:benchmarks} discusses competition
divisions and benchmarks.  The competition is being run on StarExec,
which is described in Section~\ref{sec:starexec}.
Section~\ref{sec:participants} gives an overview of participating
solvers.  Selected results are presented
%TODO: and analyzed
in Section~\ref{sec:results}.  We conclude with suggestions for future
SMT Competitions in Section~\ref{sec:conclusions}.

@@

SMT Competitions until~2014 employed a pseudo-random selection process
to further cull the number of benchmarks used in the competition.
However, a 2013 evaluation of SMT-COMP and SMT-LIB found that this
``significantly lessens the ability of a competition to determine
`best' solvers.''~\cite{CSW15} Together with the increased
computational resources available to the competition since its move to
StarExec, this finding prompted a paradigm shift.  Since 2015, the
competition is evaluating all solvers on all eligible benchmarks.
This change had an unintended side effect: the pseudo-random selection
process had been purposely constrained to choose benchmarks from
different benchmark families (that is, groups of benchmarks with
similar characteristics).  Since there are vast differences in size
between benchmark families in the SMT Library, the competition
since~2016 uses a scoring method that assigns a weight to each
main-track benchmark in order to de-emphasize large benchmark
families~\cite{rules18}.

Another paradigm shift concerns the treatment of benchmarks with
unknown status.  Because of the difficulty to judge solver responses
for such benchmarks, SMT Competitions until 2015 only used benchmarks
whose status was known (i.e., that had been solved previously by
existing SMT solvers).  In 2015, 30,717 non-incremental benchmarks
from the SMT Library were not used for the competition because their
status was unknown.  However, this approach was criticized for
favoring imitation of existing solvers over true innovation.  In~2016,
the SMT Library contained 29,724 non-incremental benchmarks with
unknown status; solver performance on these was evaluated but reported
separately.  Since 2017, all non-incremental benchmarks are eligible
for the competition's main track regardless of status, and they
directly affect the competition results.  Solver responses for such
benchmarks are assumed to be correct, unless there is a disagreement
between two or more (otherwise sound) solvers.  Benchmarks with
disagreements are removed from the competition results.  In 2016,
there were disagreements on 79 benchmarks; in 2017, only one benchmark
was removed; and in 2018, there were no disagreements between
otherwise sound solvers.

@@

Also shown in Figure~\ref{fig:job-pairs} is the total wall-clock time
(summed over all job pairs) for the competition.  This has more than
doubled, from 3.5~years in~2015 to 7.6~years in~2018.  These are
solver run-times only; time that the cluster spent on pre- and
post-processing of benchmarks and solver output or on other auxiliary
tasks is not included.  The increase in~2016 is largely due to the
inclusion of (non-incremental) benchmarks with unknown status; these
were evaluated with a wall-clock time limit of 10~minutes.  In~2017,
when these benchmarks had become eligible for the main track,
main-track job pairs were first run with a time limit of 10~minutes.
When it became apparent to the organizers that there were sufficient
resources to increase this limit, job pairs that had timed out were
re-run with a time limit of 20~minutes.  (In this way, about 2.9 years
of wall-clock time in 2017 were spent on job pairs that were then
re-run.  This time is not included in Figure~\ref{fig:job-pairs}.)
Table~\ref{table:time-limits} shows the wall-clock time limits that
were imposed on job pairs for each track and year.

@@

\begin{table}
  \caption{Participants (2015--2018)}
  \label{table:participants}
  \centering
  \begin{tabular}{|l|p{11.5cm}|}
    \hline
    Solver & Submitters (Affiliations)\\
    \hline
    ABC~\cite{DBLP:conf/cav/BraytonM10} & Valeriy Balabanov
                        (Mentor Graphics)\\
    Alt-Ergo          & Albin Coquereau
                        (CNRS and University of Paris-Sud)\\
    AProVE~\cite{DBLP:conf/cade/GieslBEFFOPSSST14} & Carsten Fuhs
                        (University College London; Birkbeck, University of London)\\
    Boolector~\cite{DBLP:conf/cav/NiemetzPWB18} & Mathias Preiner
                        (Johannes Kepler University)\\
    COLIBRI~\cite{colibri} & Fran\c{c}ois Bobot
                        (French Alternative Energies and Atomic Energy Commission)\\
    Ctrl-Ergo         & Mohamed Iguernlala
                        (CNRS and University of Paris-Sud)\\
    CVC3~\cite{DBLP:conf/cav/BarrettT07} & Kshitij Bansal
                        (New York University)\\
    CVC4~\cite{DBLP:conf/cav/BarrettCDHJKRT11}  & Kshitij Bansal, Clark Barrett, Andres N\"otzli, Duligur Ibeling
                        (New York University; Stanford University)\\
    MapleSTP          & Jimmy Liang
                        (University of Waterloo)\\
    MinkeyRink        & Trevor Hansen
                        (University of Melbourne)\\
    OpenSMT2~\cite{DBLP:conf/sat/HyvarinenMAS16} & Matteo Marescotti, Antti Hyv\"arinen
                        (University of Lugano)\\
    ProB~\cite{doi:10.1002/9781119002727.ch14} & Sebastian Krings
                        (University of D\"usseldorf)\\
    Q3B               & Martin Jon\'a\v{s}
                        (Masaryk University)\\
    raSAT~\cite{rasat} & Tung Vu Xuan
                        (Japan Advanced Institute of Science and Technology)\\
    Redlog~\cite{DBLP:journals/cca/Dolzmann097} & Haniel Barbosa
                        (Inria Nancy)\\
    SMT-RAT~\cite{DBLP:conf/sat/CorziliusKJSA15} & Florian Corzilius, Gereon Kremer
                        (RWTH Aachen University)\\
    SMTInterpol~\cite{DBLP:conf/spin/ChristHN12} & J\"urgen Christ
                        (University of Freiburg)\\
    SPASS-SATT         & Martin Bromberger
                        (Max Planck Institute for Informatics)\\
    STP~\cite{ganesh07} & Mate Soos, Trevor Hansen, Norbert Manthey
                        (National University of Singapore; University of Melbourne; Amazon Web Services)\\
    toysmt             & Masahiro Sakai
                        (Toshiba Corporation)\\
    Vampire~\cite{DBLP:conf/cav/KovacsV13} & Giles Reger
                        (University of Manchester)\\
    veriT~\cite{DBLP:conf/cade/BoutonODF09} & David D\'eharbe, Haniel Barbosa, Hans-J\"org Schurr
                        (Federal University of Rio Grande do Norte; Inria Nancy)\\
    XSat~\cite{DBLP:conf/cav/FuS16} & Martin Velez
                        (University of California, Davis)\\
    Yices2~\cite{Dutertre:cav2014} & Bruno Dutertre, Dejan Jovanovi\'c
                        (SRI International)\\
    Z3~\cite{DBLP:conf/tacas/MouraB08} & Christoph Wintersteiger, Aleksandar Zelji\'c
                        (Microsoft Research; Uppsala University)\\
    \hline
  \end{tabular}
\end{table}

@@

@@

Solvers in each division (of each track) are ranked according to a
metric that is based, first, on the number of erroneous responses
(which is usually~0, but solvers that did give erroneous responses are
ranked behind solvers that did not); second, on the number of correct
responses; third, on the wall-clock and CPU time that the solver
process consumed.  For the unsat-core track (which uses unsatisfiable
benchmarks only), the reduction in the size of the unsatisfiable core
is used instead of the number of correct responses.  A detailed
description of the scoring regulations can be found in the competition
rules~\cite{rules18}.

As discussed in Section~\ref{sec:benchmarks}, main-track benchmarks
with unknown status are removed from the competition results if two
(or more) solvers that are sound on benchmarks with known status
disagree on their result.  Otherwise, solver responses for benchmarks
with unknown status are assumed to be correct.

All solvers are run with four cores available, and the time limit~$T$
that is imposed on each job pair (see Table~\ref{table:time-limits})
is a wall-clock limit.  Thus, solvers that take advantage of
parallelism might use up to~$4T$ CPU time.  For the main track and the
unsat-core track, the competition reports parallel and sequential
performance separately.  The latter is determined by imposing a
(virtual) CPU time limit of~$T$.

@@

\TODO: non-competitive divisions
